Parsing Postfix log files

### 2007-09-24 doc/logparser.tex
Expanded the prior art section, making it a comparison rather than a critique.

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs.  This is used in the bibliography instead of \url{}.
\usepackage{url}
%\DeclareUrlCommand\urlwrap{}
%\newcommand{\urlwrapfill}[1]{\urlwrap{#1}\hfill}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}
% When creating a PDF make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author don't work - I think either gnuplot or graphviz clobbers it.
\usepackage{hyperref}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
%    breaklinks  = true,
}
\usepackage{breakurl}

@@

XXX SHOULD THERE BE A REGEX SECTION\@?

@@

There only appears to be one prior paper published about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of Log Mail Analyzer is
quite different from this parser: it attempts to present correlated log
data in a form suitable for a systems administrator to search using
grep(1), awk(1) and the myriad of other Unix utilities already available,
producing a CSV file which can subsequently be imported into a
MySQL or Berkeley DB database.  The decision to support both CSV and
Berkeley DB appears to have been a limiting factor:
Comma-Separated Value is a very simple format
where a record is stored in a single line, with fields separated by a comma
or other punctuation symbol.  Problems with CSV files include the
difficulty in escaping separators, providing multiple values for a field
(e.g.\ multiple recipients), adding new fields, and lack of a bundled
schema describing the fields.

Berkeley DB is not an SQL database, it only supports (key, value) pairs; in
the main table the value used is a CSV line, and the key is an integer
referred to by other tables.  The secondary by-sender, by-recipient,
by-date and by-IP tables use the sender/recipient/date/IP as the key, and a
CSV list of integers referring to the main table is used as the value.  In
effect this re-implements SQL foreign keys, but without the full
functionality offered by even the most basic of SQL databases.  It also
requires custom code to search on some combination of the above, though the
authors of LMA did provide some queries: IP-STORY, FROM-STORY, DAILY-EMAIL
and DAILY-REJECT\@.  Berkeley DB appears to be the least useful of the
three output formats

@@

Parsing is hand-coded and requires major changes to the code to parse new
log lines or extract additional data.  It does not appear to deal with any
of the complications discussed in
sections~\ref{complications}~and~\ref{additional complications}.  The
program fails to parse Postfix 2.2.x or 2.3.x log files; indeed it produces
a number of warnings and no usable output.

@@

XXX EXPLAIN WHY ASO IS MORE COMPREHENSIVE, WHAT IT DOES BETTER AND WHY\@.
MAKE THIS A COMPARISON INSTEAD OF A CRITIQUE\@.

>>>>

% Useful stuff for math mode.
\usepackage{amstext}
% Wrapping of URLs.  This is used in the bibliography instead of \url{}.
\usepackage{url}
%\DeclareUrlCommand\urlwrap{}
%\newcommand{\urlwrapfill}[1]{\urlwrap{#1}\hfill}
\newcommand{\parsername}{ASO }
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}
\renewcommand{\refname}{Bibliography}
% Add the bibliography into the table of contents.
\usepackage[section,numbib]{tocbibind}
% When creating a PDF make the table of contents into links to the pages
% (without horrible red borders) and include bookmarks.  The title and
% author don't work - I think either gnuplot or graphviz clobbers it.
\usepackage{hyperref}
\hypersetup{
    pdftitle    = {Parsing Postfix log files},
    pdfauthor   = {John Tobin},
    final       = true,
    pdfborder   = {0 0 0},
%    breaklinks  = true,
}
\usepackage{breakurl}

@@

@@

There only appears to be one prior paper published about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of Log Mail Analyzer is
quite different from this parser: it attempts to present correlated log
data in a form suitable for a systems administrator to search using
grep(1), awk(1) and the myriad of other standard Unix text processing
utilities already available, producing a CSV file which can subsequently be
imported into a MySQL or Berkeley DB database.  The decision to support
both CSV and Berkeley DB appears to have been a limiting factor:
Comma-Separated Value is a very simple format where a record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with CSV files include the difficulty in escaping separators,
providing multiple values for a field (e.g.\ multiple recipients), adding
new fields, and lack of a bundled schema describing the fields.

Berkeley DB is not an SQL database, supporting only (key, value) pairs; in
the main table the key is an integer referred to by secondary tables and
the value used is a CSV line containing all the data for that row.  The
secondary by-sender, by-recipient, by-date and by-IP tables use the
sender/recipient/date/IP as the key, and a CSV list of integers referring
to the main table is the value.  In effect this re-implements SQL foreign
keys, but without the functionality offered by even the most basic of SQL
databases (joins, ordering, searches, etc.).  It also requires custom code
to search on some combination of the above, though the authors of LMA did
provide some queries: IP-STORY, FROM-STORY, DAILY-EMAIL and DAILY-REJECT\@.
Berkeley DB appears to be the least useful of the three output formats

@@

\parsername is bundled with extensive documentation, including an in-depth
explanation of the database schema.  It supports multiple values for every
field for which multiple values are reasonable (e.g. the sender address
should have one value only, whereas the recipient address can have
multiple values).  The default database it uses is SQLite3, which supports
most standard SQL features (its lacks full foreign key support), has
libraries available for most common programming languages, and a simple
text based client allowing ad-hoc queries with minimal effort.

Parsing in LMA is hand-coded and requires major changes to the code to
parse new log lines or extract additional data.  It does not appear to deal
with any of the complications discussed in
sections~\ref{complications}~and~\ref{additional complications}.  The
program fails to parse Postfix 2.2.x or 2.3.x log files; indeed it produces
a number of warnings and no usable output.  Parsing in \parsername is
divided in two (see sections~\ref{rules} and ~\ref{parsing-algorithm} for a
more detailed explanation): parsing a new log line is a trivial matter of
adding a new rule, but code changes may be required if new complications
(sections~\ref{complications} and~\ref{additional complications}) must be
dealt with or a new action is required (section~\ref{actions-in-detail}).

LMA does provide some simple reports (IP-STORY, FROM-STORY, DAILY-EMAIL and
DAILY-REJECT); the author was unable to parse any logs with LMA, so didn't
have an opportunity to review the reports.  \parsername does not have any
bundled reports, it delegates that task to a separate, dedicated program.

@@
